### 2014-06-17 Manuscript/Tables/Group_results.tex
Tables init

\begin{tabular}{rllcccc}
    \hline
    set & metric & framework & parametric & $stats^1$ & df & p.value \\
    \hline
    $M_L$         & Robinson-Fould & ML       & no & 151.438  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.001278e-31
                  &                & Bayesian & no & 65.62    & 4   & \textbf{0.00} \\ 
                  & Triples        & ML       & no & 142.8329 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains  - 6.983434e-30
                  &                & Bayesian & no & 4.56     & 4   & 0.34          \\ 
    $M_F$         & Robinson-Fould & ML       & no & 179.9238 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 7.743067e-38
                  &                & Bayesian & no & 85.61    & 4   & \textbf{0.00} \\
                  & Triples        & ML       & no & 150.059  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.977386e-31
                  &                & Bayesian & no & 3.01     & 4   & 0.56          \\
    $M_C$         & Robinson-Fould & ML       & no & 151.3818 & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.029428e-31
                  &                & Bayesian & no & 64.23    & 4   & \textbf{0.00} \\
                  & Triples        & ML       & no & 138.374  & 4   & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 6.290979e-29
                  &                & Bayesian & no & 26.13    & 24  & 0.35          \\
    $M_L+M_F$     & Robinson-Fould & ML       & no & 734.9213 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.104542e-139
                  &                & Bayesian & no & 317.97   & 24  & \textbf{0.00} \\
                  & Triples        & ML       & no & 531.5558 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 4.581065e-97
                  &                & Bayesian & no & 3.01     & 24  & 0.56          \\
    $M_L+M_C$     & Robinson-Fould & ML       & no & 542.3233 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 2.619808e-99
                  &                & Bayesian & no & 290.10   & 24  & \textbf{0.00} \\
                  & Triples        & ML       & no & 437.7413 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -  1.28493e-77
                  &                & Bayesian & no & 22.19    & 24  & 0.57          \\
    $M_F+M_C$     & Robinson-Fould & ML       & no & 812.206  & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 5.463138e-156
                  &                & Bayesian & no & 385.96   & 24  & \textbf{0.00} \\
                  & Triples        & ML       & no & 528.9215 & 24  & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains - 1.619383e-96
                  &                & Bayesian & no & 20.23    & 24  & 0.68          \\ 
    $M_L+M_F+M_C$ & Robinson-Fould & ML       & no & 3603.808 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains -3.819574e-169
                  &                & Bayesian & no & 1167.40  & 124 & \textbf{0.00} \\
                  & Triples        & ML       & no & 2047.923 & 124 & \textbf{0.00} \\ %10/06 - Missing 8 ML Chains 
                  &                & Bayesian & no & 115.86   & 124 & 0.69          \\
    \hline
\end{tabular}

$^1$ F-value for parametric tests and Kruskal Wallis $chi^2$ for non parametric tests.