### 2014-12-01 Manuscript/SupplementaryMaterial/SupplementaryMaterial.tex
Updated supplementary

\subsection{Tree Building}
  \input{SupplementaryMaterial/TreeBuilding}
  \label{Supp_TreeBuilding}

\subsection{Tree Comparisons}
  \input{SupplementaryMaterial/TreeComparison}
  \label{Supp_TreeComparison}

\subsection{Codes}
All codes are available at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Functions$

The tree comparison results analysis can be repeated for more details at: $https://github.com/TGuillerme/Total\_Evidence\_Method-Missing\_data/tree/master/Analysis$ %fully running! Try it at home (5-10 minutes of data loading)!

\subsection{Additional Results}
  \input{SupplementaryMaterial/Supp_results}
  \label{Supp_results}