### 2014-12-01 Manuscript/SupplementaryMaterial/TreeComparison.tex

\section{Supplementary material Section 2}

\subsection{Triplets distance details ($T_{x,y}$)}
Triples distance ($T_{x,y}$; \citealp{dobson1975triplets}) measures the number of sub-trees made up of three taxa (triplets) that differ between two given trees.
Each triplet can be written as $I_{ijk}$=(\textit{ijk}). Where $I_{ijk}$ is equal to zero if the the two triplets (\textit{ijk}) are the same in the two trees otherwise $I_{ijk}$ is equal to one.
For any rooted binary tree there are only three possible combinations for each triplet: ((\textit{j},\textit{k}),\textit{i});, ((\textit{i},\textit{k}),\textit{j}); and ((\textit{i},\textit{j}),\textit{k}); \citep{johnson1998}.
If the trees used are not fully binary, a fourth triplet combination is possible: (\textit{i},\textit{j},\textit{k}).
We can calculate the triplet distance between two trees, $S_n$, as:
\begin{equation}
S_n = \sum_{ijk} I_{ijk}
\end{equation}
Where:
\begin{equation}
\sum_{ijk} = \binom{n}{4} = \frac{n!}{4!(n-4)!}
\end{equation}
And where \textit{n} is the total number of taxa in both trees (modified from \citet{critchlowthe1996}).
If $S_n$ = 0, the trees are identical; when $S_n$ = $\binom{n}{4}$, the trees are as different as possible (i.e. every taxon has a different placement in the two trees).
Because the possible number of triplets per clade is a finite number, the probability of two random trees with the same n taxa to have the same triplet is:
\begin{equation}
P({I_{ijk}}=0) = \frac{1}{4}
\end{equation}
Therefore one can calculate the probability of two random trees having the same triplets: 
\begin{equation}
P({S_{n}}=0) = \sum_{ijk} P_{I_{ijk}=0}
\end{equation}
\begin{equation}
P({S_{n}}=0) = \frac{n!}{4(3!(n-3)!}
\end{equation}
And in the same way:
\begin{equation}
P({S_{n}}=1) = \frac{3n!}{4(3!(n-3)!}
\end{equation}

\subsection{Robinson-Foulds distance details}
Robinson-Foulds distance (\textit{RF}; \citealp{RF1981}), or "path difference"
% NC: path difference or *distance*? - TG: difference.
, measures the number of shared clades across two trees. The metric reflects the distance between the distributions of tips among clades in the two trees \citep{RF1981} and can be expressed as following:
\begin{equation}
RF_{x,y} = N_{x} + N_{y} - 2C_{x,y}
\end{equation}
Where $C_{x,y}$ is the number of clades in common in the two trees. $C$ % NC: Why italics? Why not $C$?
is one if the two trees have the same $n$ taxa; the maximal value is $C = n-2$. % NC: again why italics for these bits? 
%NC: I don't get this. Explain more clearly. also is it important?
This metric is more sensitive to taxon displacement than Triples distance (i.e. if one taxon moves out of a clade, then the clades are no longer considered similar; \citet{critchlowthe1996,johnson1998,wiensmissing2003}).
The minimal value of $C$ is equal to 1 if the two trees have the same n taxa;
the maximal value in $C = n-2$.
For a fully unresolved tree (star tree) $N$=1 and for a fully resolved tree (binary tree) $N = n-2$.
The minimal and maximal topological distance for  taxa is:
\begin{equation}
RF_{min} = 1 + 1 - 2C_{x,y}
\end{equation}
And:
\begin{equation}
RF_{max} = 2(n-2)-2
\end{equation}
One can then rescale \textit{RF.scaled} by using the maximal and minimal value for any $n$ taxa:
\begin{equation}
RF.scaled_{x,y} = \frac{RF_{x,y}-RF_{max}}{RF_{max}}
\end{equation}
This metric is more sensitive to taxa displacement than the Triplet distance \citep{critchlowthe1996,johnson1998,wiensmissing2003} and therefore a low value will show a good clade conservation between two trees and a high value will show a bad recovery of common clades.

\subsection{Normalised Tree Similarity}
For any tree with \textit{n} taxa compared using a tree distance metric $m$, Normalized Tree Similarity, $NTS_m$ (\citealp{Bogdanowicz2012}), represents the similarity score for the two trees given the expected distance between two random Yule trees with $n$ taxa. If $\bar{d}_{m,n}$\textit{(rand)} is the average distance between two random Yule trees with $n$ taxa and $d_{m,n}$\textit{(x,y)} the distance between the two trees \textit{x} and \textit{y} containing each $n$ taxa, then:
\begin{equation}
NTS_{m,n}(x,y)=\frac{\bar{d}_{m,n}(rand) - d_{m,n}(x,y)} {\bar{d}_{m,n}(rand)}
\end{equation}
\textit{NTS} ranges from one to -$\infty$.
For any $m,n$, when $NTS$ = 1, the trees are identical, when \textit{NTS} = 0 the trees are no more different than expected by chance, and when $NTS$ $<$ 0, the trees are more different than expected when comparing two random trees.

\subsection{Tree comparisons}
\subsubsection{Random tree comparison scaling}
We used the comparison of 1000 random trees to obtain the mean comparison value $\bar{d}_{m,n}$\textit{(rand)} for the NTS metric.
We randomly generated two sets of 1000 trees of \textit{n} taxa using the rmtree function of ape package (v3.0-11 \citet{paradisape:2004}) that generates a given number of random Yule trees.
We calculated the $\bar{d}_{m,n}$\textit{(rand)} value using an approach similar to the RPCBTC (described below) by performing 1000 random pairwise comparisons using the TreeCmp java script \citep{Bogdanowicz2012}.