### 2015-01-11 Manuscript/SupplementaryMaterial/Supp_results.tex

\section{Supplementary material Section 3}

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{Figures/Supplementary/Boot+Bayt-AllParam-RF+Tr-BW.pdf}
\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood Bootstrap trees (black) and Bayesian posterior treee distribution (grey). The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Supp_BootBayt_allparam} %Differences between all the parameters and between two methods (Boot vs Bayt)
\end{figure}

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-ML-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
\caption{Effect of missing data on topological recovery using the Maximum Likelihood trees. The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis. Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
\label{Fig_Supp_paircomp_ML}
\end{figure} %Pairwise BC for the ML trees for RF and TR. (parameters differences)

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Boot-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
\caption{Effect of missing data on topological recovery using the Maximum Likelihood bootstrap trees. The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis. Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
\label{Fig_Supp_paircomp_Boot}
\end{figure} %Pairwise BC for the ML bootstrap trees for RF and TR. (parameters differences)

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/Supplementary/PPairwiseComp-Bayt-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{Figures/Supplementary/PairwiseComp-Bayt-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
\caption{Effect of missing data on topological recovery using the Bayesian posterior tree distribution. The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis. Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
\label{Fig_Supp_paircomp_Bayt}
\end{figure} %Pairwise BC for the Bayesian posterior trees distribution for RF and TR. (parameters differences)

%Summary of metrics values for all parameters combinations
\begin{table}[ht]
\label{Tab_Supp_summary_metric_allparam}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF^1$ & 0.06 & 0.26 & 0.40 & 0.41 & 0.50 & 0.95 \\ 
                                        & $Tr^2$ & 0.29 & 0.45 & 0.59 & 0.63 & 0.84 & 1.00 \\ 
  Bayesian consensus                    & $RF^1$ & 0.69 & 0.71 & 0.72 & 0.76 & 0.79 & 0.96 \\ 
                                        & $Tr^2$ & -0.28 & -0.11 & 0.17 & 0.19 & 0.37 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF^1$ & 0.06 & 0.18 & 0.27 & 0.26 & 0.34 & 0.46 \\ 
                                        & $Tr^2$ & 0.23 & 0.31 & 0.35 & 0.38 & 0.45 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF^1$ & 0.16 & 0.22 & 0.32 & 0.34 & 0.42 & 0.65 \\ 
                                        & $Tr^2$ & 0.24 & 0.35 & 0.40 & 0.50 & 0.67 & 0.98 \\ 
   \hline
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}

%Summary of metrics values for the ML parameter
\begin{table}[ht]
\label{Tab_Supp_summary_metric_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF^1$ & 0.44 & 0.51 & 0.63 & 0.66 & 0.78 & 0.95 \\ 
                                        & $Tr^2$ & 0.45 & 0.56 & 0.76 & 0.74 & 0.93 & 0.99 \\ 
  Bayesian consensus                    & $RF^1$ & 0.71 & 0.73 & 0.80 & 0.82 & 0.88 & 0.95 \\ 
                                        & $Tr^2$ & 0.37 & 0.46 & 0.67 & 0.67 & 0.87 & 0.96 \\ 
  Maximum Likelihood bootstraps         & $RF^1$ & 0.34 & 0.37 & 0.42 & 0.41 & 0.44 & 0.46 \\ 
                                        & $Tr^2$ & 0.32 & 0.40 & 0.51 & 0.46 & 0.51 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF^1$ & 0.33 & 0.41 & 0.52 & 0.50 & 0.60 & 0.65 \\ 
                                        & $Tr^2$ & 0.41 & 0.56 & 0.76 & 0.71 & 0.84 & 0.98 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}

%Summary of metrics values for the MF parameter
\begin{table}[ht]
\label{Tab_Supp_summary_metric_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF^1$ & 0.23 & 0.46 & 0.64 & 0.61 & 0.79 & 0.93 \\ 
                                        & $Tr^2$ & 0.65 & 0.84 & 0.95 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF^1$ & 0.72 & 0.77 & 0.86 & 0.85 & 0.94 & 0.96 \\ 
                                        & $Tr^2$ & -0.16 & 0.19 & 0.63 & 0.52 & 0.96 & 0.98 \\ 
  Maximum Likelihood bootstraps         & $RF^1$ & 0.14 & 0.30 & 0.40 & 0.35 & 0.45 & 0.46 \\ 
                                        & $Tr^2$ & 0.37 & 0.49 & 0.54 & 0.51 & 0.56 & 0.57 \\ 
  Bayesian posterior tree distributions & $RF^1$ & 0.24 & 0.45 & 0.57 & 0.51 & 0.63 & 0.65 \\ 
                                        & $Tr^2$ & 0.44 & 0.81 & 0.86 & 0.82 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}

%Summary of metrics values for the MC parameter
\begin{table}[ht]
\label{Tab_Supp_summary_metric_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Tree inference method & Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
  Maximum Likelihood                    & $RF^1$ & 0.40 & 0.50 & 0.64 & 0.65 & 0.79 & 0.94 \\ 
                                        & $Tr^2$ 0.70 & 0.84 & 0.93 & 0.89 & 0.99 & 1.00 \\ 
  Bayesian consensus                    & $RF^1$ & 0.76 & 0.79 & 0.86 & 0.86 & 0.92 & 0.96 \\ 
                                        & $Tr^2$ & 0.05 & 0.16 & 0.53 & 0.50 & 0.87 & 0.92 \\ 
  Maximum Likelihood bootstraps         & $RF^1$ & 0.25 & 0.34 & 0.42 & 0.38 & 0.45 & 0.46 \\ 
                                        & $Tr^2$ & 0.38 & 0.47 & 0.55 & 0.51 & 0.57 & 0.58 \\ 
  Bayesian posterior tree distributions & $RF^1$ & 0.32 & 0.44 & 0.58 & 0.52 & 0.62 & 0.65 \\ 
                                        & $Tr^2$ & 0.39 & 0.78 & 0.82 & 0.79 & 0.98 & 0.98 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}

%Summary of the BC between pairs of methods for the ML parameter
\begin{landscape}
\begin{table}[ht]
\label{Tab_Supp_summary_BC_ML}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.30 & 0.31 & 0.69 & 0.61 & 0.77 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.79 & 0.81 & 0.84 & 0.86 & 0.85 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.22 & 0.29 & 0.36 & 0.54 & 0.69 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.42 & 0.53 & 0.51 & 0.74 & 0.78 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.49 & 0.61 & 0.51 & 0.67 & 0.74 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.61 & 0.70 & 0.63 & 0.81 & 0.81 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.02 & 0.03 & 0.04 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.69 & 0.78 & 0.64 & 0.79 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.01 & 0.02 & 0.02 & 0.04 & 0.08 & 0.09 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.74 & 0.75 & 0.68 & 0.84 & 0.87 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.75 & 0.85 & 0.85 & 0.95 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.96 & 0.95 & 0.97 & 0.98 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}
\end{landscape}

%Summary of the BC between pairs of methods for the MF parameter
\begin{landscape}
\begin{table}[ht]
\label{Tab_Supp_summary_BC_MF}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.25 & 0.48 & 0.50 & 0.76 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.38 & 0.69 & 0.75 & 0.72 & 0.80 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.18 & 0.32 & 0.36 & 0.47 & 0.77 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.34 & 0.40 & 0.38 & 0.53 & 0.55 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.47 & 0.71 & 0.60 & 0.86 & 0.94 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.54 & 0.62 & 0.56 & 0.64 & 0.80 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & 0.01 & 0.01 & 0.01 & 0.03 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.54 & 0.49 & 0.70 & 0.75 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.02 & 0.02 & 0.04 & 0.04 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.29 & 0.66 & 0.54 & 0.72 & 0.82 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.69 & 0.72 & 0.71 & 0.72 & 0.72 \\ 
                                                                       & $Tr^2$ & 0.91 & 0.91 & 0.91 & 0.93 & 0.92 & 0.98 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}
\end{landscape}

%Summary of the BC between pairs of methods for the MC parameter
\begin{landscape}
\begin{table}[ht]
\label{Tab_Supp_summary_BC_MC}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\  
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.03 & 0.32 & 0.66 & 0.55 & 0.75 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.51 & 0.69 & 0.80 & 0.76 & 0.80 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.17 & 0.21 & 0.31 & 0.46 & 0.68 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.31 & 0.39 & 0.39 & 0.56 & 0.61 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.44 & 0.47 & 0.52 & 0.78 & 0.90 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.52 & 0.59 & 0.55 & 0.66 & 0.77 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.01 & 0.01 & 0.02 & 0.02 & 0.03 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.47 & 0.62 & 0.51 & 0.66 & 0.73 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.02 & 0.04 & 0.04 & 0.05 & 0.06 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.45 & 0.64 & 0.57 & 0.74 & 0.79 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.69 & 0.73 & 0.73 & 0.76 & 0.81 & 0.86 \\ 
                                                                       & $Tr^2$ & 0.91 & 0.92 & 0.93 & 0.94 & 0.96 & 0.99 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\end{table}
\end{landscape}