### 2015-01-25 Manuscript/SupplementaryMaterial/SupplementaryMaterial.tex

\section{Code}
All code for performing the analyses is available at: \url{https://github.com/TGuillerme/Total_Evidence_Method-Missing_data}.

\newpage
\section{Supplementary Material 1: Tree Building}
  \input{SupplementaryMaterial/TreeBuilding}
  \label{Supp_TreeBuilding}

\newpage
\section{Supplementary Material 2: Tree Comparisons}
  \input{SupplementaryMaterial/TreeComparison}
  \label{Supp_TreeComparison}

\newpage
\section{Supplementary Material 3: Additional Results}
  \input{SupplementaryMaterial/Supp_results}
  \label{Supp_results}

%\newpage
%\section{Supplementary References}
%    \bibliographystyle{sysbio}
%    \bibliography{SupplementaryMaterial/Supp_Ref}